The Relationship between Family Status and Academic Achievement in Children

### 2023-03-24 outputs/paper/paper.qmd
second version of the draft

You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.

In the past, family status variables, such as the socio-economic status and educational level of parents, were commonly considered as strong predictors of academic achievement in children. More specifically, parents who have lower levels of education, those with higher levels of education are more inclined to view higher education as desirable and encourage their children to excel academically. They also tend to hold higher expectations for their children's academic performance. 
(https://www.academia.edu/2712079/How_does_parents_education_level_influence_parenting_and_childrens_achievement)

One potential explanation for this relationship is that higher levels of education can provide parents with access to resources such as income, time, energy, and community contacts, which can enable greater involvement in a child's education. As a result, the influence of family status variables on children's academic achievement may be best understood as a complex interaction between status and process variables (https://files.eric.ed.gov/fulltext/EJ1079955.pdf).

However, recent research has indicated that the relationship between these factors and academic achievement is not always direct. Instead, socio-economic status and parents' education are part of a larger set of psychological and sociological variables that can impact children's educational outcomes (https://files.eric.ed.gov/fulltext/EJ1079955.pdf).

This paper will test and explore how parents’ social-economics status can affect childrens’ education. The data that will be used in this paper comes from the US General Social Survey from National Opinion Research Center at the University of Chicago. The factor of socio-economics status is measured using the occupation prestige score. It is a measure used in social science research to assess the level of social status or prestige associated with a particular occupation or profession, which were developed by the National Opinion Research Center at the University of Chicago. As part of the GSS, respondents are asked to rate the prestige or social standing of various occupations on a scale of 1 to 90, with higher scores indicating greater prestige. Another variable that has been used is the education level, which is a number of variables to indicate the number of years the respondents have spent in school and college.
(https://gss.norc.org/Documents/reports/methodological-reports/MR122%20Occupational%20Prestige.pdf)

*MISSING*briefly discuss the results and application.

@@

Our data is of penguins (@fig-bills).

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-degree
#| tbl-cap: Respondents' Degree in 2021 Survey

tb1 <- cleaned_data %>% 
  count(degree) %>% 
  mutate(percent=percent(x=n/sum(n))) %>% 
  rename(Degree = degree, Frequency = n, Proportion = percent)

tb1<-tb1%>%arrange(factor(tb1$Degree,c("graduate", "bachelor's", "associate/junior college", "high school", "less than high school", NA)))

tb1%>%kable()
```

@@

  # geom_histogram(bins = 15, fill = "lightblue", color = "white") +
  labs(x = "Education (in Years)", y = "Desensity") +
  scale_x_continuous(breaks = seq(0, 20, by = 1)) +
  theme_bw()
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-educationyears
#| tbl-cap: Number of respondents by gender for 2021 survey
tb0 <- cleaned_data %>% summarise(Min = min(cleaned_data$educ,na.rm=T),
                              Median = median(cleaned_data$educ,na.rm=T),
                              Max = max(cleaned_data$educ,na.rm = T),
                              Mean = mean(cleaned_data$educ,na.rm=T))

tb0 %>% kable()
```

@@

#prestige score distribution
ggplot(cleaned_data, aes(x = mapres10)) +
  geom_histogram(aes(y = ..density.., fill = "Mother's Prestige Score"), alpha = 1, binwidth = 1, color = "black")+
  geom_histogram(aes(x = papres10, y = ..density.., fill = "Father's Prestige Score"), alpha = 0.8, binwidth = 1, color = "black") +
  scale_fill_manual(values = c("Mother's Prestige Score" = "darkgreen", "Father's Prestige Score" = "lightgreen"), name = "Variables") +
  labs(x = "Prestige Score",
       y = "Density") +
  theme_bw()
```

```{r}
#| label: fig-presandeduc
#| fig-cap: Mother's and Father's Prestige Score
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 4
#pres vs. educ
p1 <- ggplot(cleaned_data, aes(x=educ, y=papres10)) +
  geom_point(color="#69b3a2", size=2,alpha=0.3) +
  labs(x="Education (in Years)", y="Father's Prestige Score")+
  theme_bw()
p2 <- ggplot(cleaned_data, aes(x=educ, y=mapres10)) +
  geom_point(color="darkgreen",size=2,alpha=0.3) +
  labs(x="Education (in Years)", y="Mother's Prestige Score")+
  theme_bw()

p1 + p2
```

```{r}
#| label: fig-parentseduc
#| fig-cap: Mother's Education and Father's Education Level
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 4
ggplot(cleaned_data, aes(x = maeduc)) +
geom_histogram(aes(y = ..density.., fill = "Mother's Education"), alpha = 1, binwidth = 0.5, color = "black") +
geom_histogram(aes(x = paeduc, y = ..density.., fill = "Father's Education"), alpha = 0.8, binwidth = 0.5, color = "black") +
scale_fill_manual(values = c("Mother's Education" = "darkgreen", "Father's Education" = "lightgreen"), name = "Variables") +
labs(x = "Education (in Years)",
y = "Density") +
theme_bw()
```

Talk more about it.

Also bills and their average (@fig-billssssss). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)

```{r}
#| label: fig-birth
#| fig-cap: Birthplace of the Respondent (if outside of US)
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 4

ggplot(cleaned_data, aes(x = educ, fill = factor(born))) +
  geom_histogram(position = "dodge", binwidth = 1.5, alpha = 0.5, color = "black")+
   labs(x = "Education (in Years)",
       y = "Count",
       fill = "Birthplace of the Respondent (if in, of US)")+
  theme_bw()
```
```{r,echo=FALSE}
tb3 <- cleaned_data %>% 
  count(born) %>% 
  mutate(percent=percent(x=n/sum(n)))
```

```{r}
#| label: fig-region
#| fig-cap: Respondent's Living Region at Age of 16 v.s. Education
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 4
ggplot(cleaned_data, aes(x = educ,fill = factor(reg16))) +
  geom_histogram(binwidth = 2, color = "black") +
  facet_wrap(~ reg16, ncol = 2)+
  labs(x = "Education (in Years)",
       y = "Count")+
  theme_bw()+
  theme(legend.position = "none")
```

@@

@@

ggplot(cleaned_data, aes(x = educ, y = paeduc, color = incom16)) +
  geom_jitter(alpha = 0.2) +
  stat_smooth(method = "lm", se = FALSE) +
  facet_wrap(~incom16)+
  labs(x = "Education (in Years)",
       y = "Father's Educaion (in Years)",
  color = "Family Income (at Age of 16)") +
  scale_color_discrete(name = "Family Income (at Age of 16)") +
  theme_bw()
```

@@

modelsummary::modelsummary(
  list(
    "Children's Education" = model
  ),
  fmt = 2,
  statistic = "conf.int"
)
```

Talk way more about it.

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$ {#eq-bayes}

@eq-bayes seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

## Weaknesses and next steps {#sec-weaknesses}
*problems with prestige score:*

Subjectivity: Even measures of prestige that attempt to capture objective indicators of social status, such as income or educational attainment, are still subject to subjective interpretation. People's perceptions of what constitutes a high-prestige occupation may vary based on their own cultural, social, or personal biases.

Historical and cultural context: Measures of prestige are not fixed over time and may vary across different societies or historical periods. For example, certain occupations that were highly prestigious in the past, such as blacksmith or scribe, may be less highly regarded today.

Heterogeneity within occupations: Different individuals within the same occupation may have different levels of prestige, depending on factors such as their level of experience, education, or specialization.

Overemphasis on certain aspects of occupation: Some measures of prestige may place too much emphasis on certain aspects of occupation, such as income or educational requirements, while overlooking other important factors such as job security, working conditions, or level of autonomy.

Lack of nuance: Prestige scores may provide a broad measure of social status and occupational major groups, but they may not capture the complexity and diversity of occupational experiences, especially for those in lower-prestige occupations.
(https://www.researchgate.net/publication/236747304_A_Scale_of_Occupational_Prestige_in_Canada_Based_on_NOC_Major_Groups)

*problems with missing data:*

Reduced statistical power is one such issue, which lowers the likelihood of rejecting the null hypothesis when it is false.
Missing data can also cause bias in parameter estimation.
The representativeness of the sample can also be reduced due to missing data.
The analysis of the study can become more complex when missing data is present.
These problems can threaten the validity of the trials and lead to invalid conclusions.
(https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3668100/)

>>>>

In the past, family status variables, such as the socio-economic status and educational level of parents, were commonly considered as strong predictors of academic achievement in children. More specifically, parents who have lower levels of education, and those with higher levels of education are more inclined to view higher education as desirable and encourage their children to excel academically. They also tend to hold higher expectations for their children's academic performance [@davis_2014].

One potential explanation for this relationship is that higher levels of education can provide parents with access to resources such as income, time, energy, and community contacts, which can enable greater involvement in a child's education. As a result, the influence of family status variables on children's academic achievement may be best understood as a complex interaction between status and process variables [@khan].

However, recent research has indicated that the relationship between these factors and academic achievement is not always direct. Instead, socio-economic status and parents' education are part of a larger set of psychological and sociological variables that can impact children's educational outcomes [@khan].

In short, the findings in this paper demonstrate that parents’ education has a positive correlation with children’s education. Additionally, the region of living during the teenage years can also impact children’s education. Parents’ prestige score seems to be another factor in children’s education. However, there is some limitation on the definition of prestige score that may influence the overall results. This will be discussed in the @sec-weakness.

The data will be presented clearly and succinctly with plots and tables. This following analysis is processed in R [@R]  with packages of tidyverse [@tidyverse], dplyr [@dplyr], here [@here], haven [@haven] and broom [@broom]. The tables are constructed via knitr [@knitr], scales [@scales] and kableExtra [@kableExtra]. The package inside tidyverse helps to create the plots in ggplot2 [@tidyverse]. This paper is knitted as a PDF file by the packages of R markdown [@RMarkdown] and formatted using patchwork [@patchwork].

@@

## Source
The 2021 General Social Survey (GSS) is a nationally representative survey conducted to collect data on social trends and attitudes among people living in the United States. Data was collected through face-to-face interviews with adult residents, covering a wide range of variables that are of interest to social scientists, policymakers, and the general public. The variables measured in the survey include demographics, employment, education, health, family, and social attitudes [@gss_data_explorer].

The target population is US adults aged 18 and older. The GSS uses a multistage probability sampling approach to recruit its sample from the US Census Bureau's Master Address File. One strength of the GSS is its long history and large sample size, but a potential weakness is a reliance on self-reported data, which can be subject to response bias. The questionnaire includes both closed-ended and open-ended questions, but there may be concerns about biased or leading questions. The GSS employs various methods to adjust for non-response and ensure the representativeness of the sample. One strength of the GSS is its long history and large sample size, but a potential weakness is a reliance on self-reported data, which can be subject to response bias. The questionnaire includes both closed-ended and open-ended questions, but there may be concerns about biased or leading questions. [@norcuofch] The GSS employs various methods to adjust for non-response and ensure the representativeness of the sample. The limitation will be explained in detail in the @sec-weakness.

This paper will test and explore how parents’ social-economics status can affect children’s education. The data that will be used in this paper comes from the US General Social Survey from the National Opinion Research Center at the University of Chicago. The factor of socio-economics status is measured using the occupation prestige score. It is a measure used in social science research to assess the level of social status or prestige associated with a particular occupation or profession, which was developed by the National Opinion Research Center at the University of Chicago. As part of the GSS, respondents are asked to rate the prestige or social standing of various occupations on a scale of 1 to 90, with higher scores indicating greater prestige. [@smith] Another variable that has been used is the education level, which is a number of variables to indicate the number of years the respondents have spent in school and college. Other variables will be discussed in @sec-data.

## Data cleaning

For simplicity purposes, only variables that are closely related to the topics are selected for further analysis. These variables are mapres10, papres10, paeduc, maeduc, educ, mawrkslf, pawrkslf, degree, born, sex, reg16, and incom16.

- **mapres10**: A numeric variable indicating the respondent's mother’s prestige score.
- **papres10**: A numeric variable indicating the respondent's father’s prestige score.
- **paeduc**: A numeric variable indicating the number of years of education completed by the respondent's father.
- **maeduc**: A numeric variable indicating the number of years of education completed by the respondent's mother.
- **educ**: A numeric variable indicating the number of years of education completed by the respondent.
- **mawrkslf**: A categorical variable indicating whether the respondent's mother was employed for wages or salary.
- **pawrkslf**: A categorical variable indicating whether the respondent's father was employed for wages or salary.
- **degree**: A categorical variable indicating the highest degree earned by the respondent.
- **born**: A categorical variable indicating the region of the US where the respondent was born.
- **sex**: A categorical variable indicating the respondent's sex.
- **reg16**: A categorical variable indicating the region of the US where the respondent lived at age of 16.
- **incom16**: A categorical variable indicating the respondent's total household at age of 16.

The level of “no formal schooling” in **maeduc**, **paeduc** and **educ** is modified as the integer zero to make the variables consistent. All the numeric variables are shown as integers, but their class are in terms of character in R. Hence, the variables of **paeduc**, **maeduc**, **educ**, **mapres10** and **papres10** have been converted into a class of numeric for further analysis.

None of the missing values (NA) is removed in this paper. The reason is that removing the missing values in every column would result in a huge loss in the dataset. To maintain the representativeness of the data, missing values are automatically filtered by R, rather than removed manually. The possible limitation and effect of the missing values will be discussed in the @sec-weakness.

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-degree
#| tbl-cap: Number of respondents by degree for 2021 survey
tb1 <- cleaned_data %>% count(degree) %>%
  mutate(percent=percent(x=n/sum(n)))
tb1<-tb1%>%arrange(factor(tb1$degree,c("graduate", "bachelor's", "associate/junior college", "high school", "less than high school", NA)))
colnames(tb1) <- c("Degree", "Total", "Proportion")
tb1%>%kable(booktabs = TRUE)
```
The respondents in this study are the children and their parents' education has also been collected in the GSS. As for the response variables, children’s education level, there are two ways to measure. One is the degree, and the other one is the number of years in school. In @tbl-degree, it shows the distribution of education in terms of degrees. The data reveals that the most common degree category among the group is high school, with 1597 individuals (39.6%) having this level of education. Following this, the next most common category is a bachelor's degree, with 1036 individuals (25.7%) having completed this level of education. The data also indicates that a graduate degree is the third most common category, with 760 individuals (18.8%) having this level of education. Associate/junior college degree is less common than a bachelor's or graduate degree, with 370 individuals (9.2%) having completed this level of education. However, a smaller proportion of individuals (6.1%) have less than a high school education. Additionally, there are also 23 individuals (0.6%) whose educational degree information is not available (NA). Overall, the data provide insights into the educational distribution of the given group and may be useful for drawing conclusions and making informed decisions related to education.

@@

  # geom_histogram(bins = 15, fill = "lightblue", color = "white") +
  labs(x = "Education (in Years)", y = "Desensity") +
  scale_x_continuous(breaks = seq(0, 20, by = 1)) +
  theme_bw()
```
The second variable utilized to quantify education is relation to the number of years of schooling. In @fig-educ, this histogram shows an overall distribution of the respondents' education in years. It is clear that the distribution is right-skewed, which indicates that there are more respondents with lower levels of education than with higher levels of education. In other words, the bulk of the respondents are clustered towards the left side of the histogram (lower levels of education), and there are relatively fewer respondents towards the right side of the histogram (higher levels of education). In short, the majority of respondents in 2021 GSS are highly educated.

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-educ
#| tbl-cap: Number of respondents by education for 2021 survey
tb0 <- cleaned_data %>% summarise(Min = min(cleaned_data$educ,na.rm=T),
                              Median = median(cleaned_data$educ,na.rm=T),
                              Max = max(cleaned_data$educ,na.rm = T),
                              Mean = mean(cleaned_data$educ,na.rm=T))

tb0 %>% kable()
```
To see a more accurate numerical summary, @tbl-educ is created. In the dataset labelled @sec-data, the category denoting "no formal schooling" has been redefined as an integer value of 0. According to the information presented in the @tbl-educ table, the range of years of schooling reported by respondents spans from 0 to 20 years. The median number of years of schooling reported is 15, and the mean is 14.76904. This mean value roughly corresponds to three years in college. Taken together, these findings suggest that the respondents surveyed in the 2021 GSS have generally completed their high school education and are highly educated.

@@

#prestige score distribution
ggplot(cleaned_data, aes(x = mapres10)) +
  geom_histogram(aes(y = ..density.., fill = "Mother's Prestige Score"), alpha = 1, binwidth = 1, color = "black")+
  geom_histogram(aes(x = papres10, y = ..density.., fill = "Father's Prestige Score"), alpha = 0.8, binwidth = 1, color = "black") +
  scale_fill_manual(values = c("Mother's Prestige Score" = "darkgreen", "Father's Prestige Score" = "lightgreen"), name = "Variables") +
  labs(x = "Prestige Score",
       y = "Density") +
  theme_bw()
```
In @fig-parentpre, it shows the distribution of respondent’s father’s and mother’s prestige scores. The distribution of father’s prestige score and the mother’s prestige score have quite different distributions.

```{r}
#| label: fig-presandeduc
#| fig-cap: Mother's and Father's Prestige Score with children's education
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 4
#pres vs. educ
p1 <- ggplot(cleaned_data, aes(x=educ, y=papres10)) +
  geom_point(color="#69b3a2", size=2,alpha=0.3) +
  labs(x="Education (in Years)", y="Father's Prestige Score")+
  theme_bw()
p2 <- ggplot(cleaned_data, aes(x=educ, y=mapres10)) +
  geom_point(color="darkgreen",size=2,alpha=0.3) +
  labs(x="Education (in Years)", y="Mother's Prestige Score")+
  theme_bw()

p1 + p2
```
In order to examine the relationship between parents’ prestige scores and children’s education level. The next plot, @fig-presandeduc, is used to show the relation. These two scatter plots show the father’s prestige score and the mother’s prestige score with their children’s education. It is hard to see a clear relation or pattern based on these two scatter plots. To verify if the prestige score is one of the factors, a linear model will be used in the @sec-model.

```{r}
#| label: fig-parentseduc
#| fig-cap: Mother's Education and Father's Education Level
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 4
ggplot(cleaned_data, aes(x = maeduc)) +
geom_histogram(aes(y = ..density.., fill = "Mother's Education"), alpha = 1, binwidth = 0.5, color = "black") +
geom_histogram(aes(x = paeduc, y = ..density.., fill = "Father's Education"), alpha = 0.8, binwidth = 0.5, color = "black") +
scale_fill_manual(values = c("Mother's Education" = "darkgreen", "Father's Education" = "lightgreen"), name = "Variables") +
labs(x = "Education (in Years)",
y = "Density") +
theme_bw()
```
Another key variable is the parent’s education level. In @fig-parentseduc, it shows the distribution of the mother’s education and the father’s education. One interesting result is that the distribution of education for mother and father is similar, or even almost identical. This indicates that there is a strong correlation between the education levels of mothers and fathers. In other words, if a mother has a high level of education, it is likely that the father also has a high level of education, and vice versa.

```{r}
#| label: fig-birth
#| fig-cap: Birthplace of the Respondent (if in US)
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 4

ggplot(cleaned_data, aes(x = educ, fill = factor(born))) +
  geom_histogram(position = "dodge", binwidth = 1.5, alpha = 0.5, color = "black")+
   labs(x = "Education (in Years)",
       y = "Count",
       fill = "Birthplace of the Respondent (if in US)")+
  theme_bw()
```
Based on [@overman], the birthplace may have an impact on children’s education. In @fig-birth, it is used to test the relationship between birthplace and children’s education. Note that the question of birthplace only asks if the respondent was born in the US or not. In general, most of the respondents have received at least 10 years of education despite their birthplace. However, the bar for respondents who were born outside of the US is much higher than the other groups. This is because 87.2% of the respondents were born in the US, and only 11.0% answered “no”. Additionally, there are 72 NA, which take a proportion of 1.8%. Given the huge differences in the sample sizes between groups, it is hard to draw a conclusion of whether birthplace may affect the education level.

```{r,echo=FALSE}
tb3 <- cleaned_data %>% 
  count(born) %>% 
  mutate(percent=percent(x=n/sum(n)))
```

```{r}
#| label: fig-region
#| fig-cap: Respondent's Living Region at Age of 16 v.s. Education
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 4
ggplot(cleaned_data, aes(x = educ,fill = factor(reg16))) +
  geom_histogram(binwidth = 2, color = "black") +
  facet_wrap(~ reg16, ncol = 2)+
  labs(x = "Education (in Years)",
       y = "Count")+
  theme_bw()+
  theme(legend.position = "none")
```
According to the study [@nieuwenhuis], there appears to be an association between neighborhood and education. In the 2021 GSS, the neighbourhood variable is being measured by the living region at the age of 16. The distribution for all regions, as shown in @fig-region, follows a similar bimodal shape, with the exception of the NA group. This could indicate that the region of living has a minimal or no impact on children's education levels. However, it is important to note that the measurement of the neighborhood through living regions at age 16 may not fully capture the impact of the neighborhood on education, which will be discussed in the @sec-weakness.

@@

To examine the relationship between family income and education, the @fig-maincome and @fig-paincome visualize this relation. In @fig-maincome, the vertical axis represents the mother’s education and the horizontal axis represents the children’s education. The respondents were asked to rate their family income at age of 16 in the following category: (1) far above average, (2) above average, (3) average, (4) below average and (5) far below average. Due to missing values, the sixth group is NA, which is coloured grey. Overall, mothers’ education and children’s education are positively correlated despite the level of income. To be more specific, for respondents whose family income was far above average, the slope seems to be steeper than the other groups, which indicates that for those with higher family income, there may be a stronger relationship between the education of the mother and the education of the children.

@@

ggplot(cleaned_data, aes(x = educ, y = paeduc, color = incom16)) +
  geom_jitter(alpha = 0.2) +
  stat_smooth(method = "lm", se = FALSE) +
  facet_wrap(~incom16)+
  labs(x = "Education (in Years)",
       y = "Father's Educaion (in Years)",
  color = "Family Income (at Age of 16)") +
  scale_color_discrete(name = "Family Income (at Age of 16)") +
  theme_bw()
```
In @fig-paincome, the income groups are the same as before. However, the plot is now examining the correlation between the father’s education and the children’s education. Once again, there is a positive correlation between the education level of fathers and their children's education, regardless of family income. However, for respondents who reported a "far above average" family income, the relationship between the father's education and children's education appears to be stronger than for other income groups. This suggests that higher family income may be associated with a more pronounced link between paternal and children's education levels. Comparing the two plots, the slopes in @fig-paincome for all income groups are much steeper than the slopes in @fig-maincome. This may suggest that there is a stronger relationship between fathers' education and children's education compared to mothers' education and children's education given the same income level.

$$
\hat{Y}=\hat{\beta_0}+\hat{\beta_1}X_{mapres10}+\hat{\beta_2}X_{papres10}+\hat{\beta_3}X_{paeduc}+\hat{\beta_4}X_{maeduc}
$$ {#eq-model}

@@

modelsummary::modelsummary(
  list(
    "Children's Education" = model
  ),
  fmt = 2,
  statistic = "conf.int"
)
```
The @tbl-lmeduc is built based on the @eq-model:

- The intercept is 11.32, which means that when all the independent variables are zero, the dependent variable has a value of 11.32.
- The coefficients for the independent variables mapres10 and papres10 are both 0.01, which means that a one-unit increase in either of these variables is associated with an increase of 0.01 in the dependent variable.
- The coefficients for the independent variables paeduc and maeduc are 0.18 and 0.05, respectively. This means that a one-unit increase in paeduc is associated with an increase of 0.18 in the dependent variable, and a one-unit increase in maeduc is associated with an increase of 0.05 in the dependent variable.
- The number of observations used in the model is 1980.
The R-squared value is 0.124, which means that the independent variables explain 12.4% of the variation in the dependent variable.
- The adjusted R-squared value is 0.122.
- The AIC and BIC values are measures of the goodness of fit of the model. Lower values indicate a better fit.
- The log-likelihood is a measure of how well the model fits the data.
- The RMSE value is 2.44, which is the root mean squared error of the model. It is a measure of how much the predictions of the model deviate from the actual values.

To put it another way, there exists a positive correlation between the educational attainment and societal status of parents with that of their children. However, when examining the specific factors of fathers' education and prestige score, it appears that the correlation with children's education is stronger than that of the mothers' education or prestige score.

This suggests that fathers' educational background may have a particularly strong influence on their children's academic achievement and future opportunities. While the influence of mothers' education and social status is still significant, the impact of fathers' education appears to be even greater in shaping their children's educational outcomes. These findings highlight the importance of parental involvement and investment in their children's education, particularly that of the father's role in this regard.

@@

## Weaknesses and next steps {#sec-weakness}

As with any research, this study has its limitations and weaknesses. One significant weakness is the use of prestige scores as a variable to measure social status and occupational standing. While attempts were made to capture objective indicators of status, such as income and education, prestige scores are still subject to subjective interpretation. Cultural, social, or personal biases may influence people's perceptions of what constitutes a high-prestige occupation. Furthermore, measures of prestige are not fixed over time and may vary across different societies and historical periods. Heterogeneity within occupations can also lead to different levels of prestige among individuals. Additionally, some measures of prestige may overemphasize certain aspects of an occupation, such as income or education, while overlooking other critical factors like job security or working conditions [@prestige].

Another weakness of this study is missing data. While the decision to retain all missing values was made intentionally to avoid issues with representativeness, missing data can still affect the statistical power and accuracy of estimates. It can also reduce the representativeness of the sample, potentially limiting the generalizability of the findings. Additionally, missing data can make the analysis of the study more complex, requiring more sophisticated techniques that may increase the risk of errors [@kang].

Moreover, this study did not use hypothesis testing, which could potentially overlook some effects that might be present in the data. Therefore, the generalizability of the findings should be interpreted with caution. Future studies should aim to address these limitations and weaknesses to ensure more accurate and comprehensive research on the topic.

To shed light in future studies, there are some potential improvements. Firstly, it may be beneficial to incorporate additional measurements for social-economic statuses, such as the International Socio-Economic Index of Occupational Status (ISEI). This index is derived from the International Standard Classification of Occupations (ISCO) and comprises comparably coded data on education, occupation, and income from 73,901 full-time employed workers across 16 countries [@international].

Furthermore, new and accurate methodologies could be implemented for data analysis. It is evident that the sample sizes across the groups in @fig-birth are substantially different. Therefore, post-stratification may be a suitable technique to adjust for these discrepancies and verify the results. Moreover, in terms of the statistical model used in the study, a simple linear model was employed with only a few predictors. Future studies could build a more sophisticated model that accounts for potential interaction or quartic terms in the analysis to improve the understanding of the relationship under investigation.

In @fig-region, the results seem to be insignificant. One way to improve is to consider the measurement of the neighbourhood using living regions at age 16 may not provide a comprehensive representation of the impact of the neighbourhood on education. Other factors, such as the quality of schools and access to resources, may also play a significant role in educational outcomes but are not captured by this measure. Therefore, it is possible that the relationship between neighbourhood and education is more complex than what can be inferred from the living region at age 16 alone. While the distribution of education levels among neighbourhoods may appear similar based on @fig-region, it is still possible that there are significant differences in educational outcomes that are not apparent in the bimodal shape of the distribution. Further research using more detailed measures of neighbourhood characteristics and educational outcomes could provide a more nuanced understanding of the relationship between neighbourhood and education.

While this data and study provide valuable insights into children's education and performance, it is important to acknowledge that some key factors have not been included. One such variable is the positive involvement of parents in their children's education. Research has consistently shown that parental involvement can have a significant impact on children's academic success, yet this aspect was not measured in the 2021 GSS [@parentinvolvement]. Therefore, it is essential to include these factors in future studies to fully understand the complex interplay between different variables and their effects on children's education. By including factors like parental involvement, future studies could provide a more comprehensive understanding of the various factors that influence children's education and help to identify effective strategies for improving educational outcomes.
